\chapter*{Introduction}
\label{cha:Introduction}
\addcontentsline{toc}{chapter}{\nameref{cha:Introduction}}

\section*{Context}
\label{sec:Context}
\addcontentsline{toc}{section}{\nameref{sec:Context}}

This research work is part of the RADIAL (RADiomics Image Analysis of the Liver) project, carried out by the
institute for minimally invasive surgery from Strasbourg (IHU), where
one of the objectives is to improve the treatment and the prevention of
liver cancer.\\
Being the second leading cause of cancer-related death worldwide, liver
cancer is considered as a major public health challenge. Current ways to characterize the liver cancer are either invasive, 
through the extraction of tissues samples or non-invasive, typically 
through interpretation of medical images.\\
%radiomics
Radiomics is a new field developed in the early 2010s to extract the encoded information within medical images in order to provide a more objective and image-guided assessment of the disease.\\
Conventional radiomics however suffers from limitations related to the hand-crafted design of the extracted features or the computation of characteristics only in a manually defined region or volume of interest. Deep learning changed the way to apprehend several medical imaging related problematics thanks to its ability to encode
morphological properties in images only by using the pixel intensities
as input. The radiomics field was impacted by this new set of algorithms, and a new branch of radiomics emerged, where the delineation of the region of interest and/or the extraction of features are performed using deep learning approaches.

\section*{Outline of the thesis}
\label{sec:Outline}
\addcontentsline{toc}{section}{\nameref{sec:Outline}}

%Context
The first chapter aims at presenting the different types of liver cancers, and particularly the
hepatocellular carcinoma (\ac{hcc}), which will be the main topic of our work. We describe the different steps of hepatocarcinogenesis, which
transforms healthy hepatic tissues into cancerous ones, in terms of the
evolution of the cells and the associated pathological changes. 
We then detail the current ways to establish the diagnosis either through extraction and inspection of tissue samples (biopsy) or 
non-invasively using medical imaging techniques.  We outline the different modalities usually exploited to establish the diagnosis, and focus on X-ray computed tomography (\ac{ct}) for the rest of this
research work. \\
%Radiomics
The analysis of medical images is commonly performed by the naked eye,
but technological advances allow the creation of computer assisted
diagnosis tools to investigate the link between imaging features 
and biological characteristics.
We present in the second chapter the radiomics technique allowing the extraction of quantitative features from images. First, we introduce the conventional pipeline based on engineered features, generally referred as \ac{hcr} (Hand-Crafted Radiomics). To emphasize the limitations of this approach and to question its reproducibility across clinical studies, we review HCR-liver related paper and assess the experimental design according to the Radiomics Quality Score (RQS). Then, we introduce the alternative paradigm brought by Deep-Learning Radiomics and review its application in the context of HCC studies.\\
We then focus on a key step of the radiomics pipeline, being the automatic segmentation of liver and its tumors.
The first methods developed to perform this task were based on
anatomical prior-knowledge or manual interactions, and often failed due to  the limited amount of training samples, or when
facing pathological cases. Automatic deep learning segmentation can address most of these issues by looking for relevant features directly from the images only. We review in chapter 3 both DL-based and non DL-based liver tumor segmentation methods.\\
In chapter 4, we present our multiphase cascaded architecture to automatically segment both the liver and tumors and to partition the tumor into necrotic and active tissues.\\
In the last chapter we present our preliminary work to automatically predict the histological grade of HCCs.
Finally, conclusion and perspectives are given.

%\\
%%Deep Learning + DLR
%The emergence of deep learning (\ac{dl}) allows creation of algorithms where features are learned directly from the raw images without extra information. The radiomics field benefited from the development of DL, and a new branch was created, the \emph{\ac{dlr}} (Deep-Learning Radiomics). We review the DLR studies performed on the \ac{hcc}, and raise the issue of segmentation, often manually performed and thus suffering from high inter and intra-variability. \\
%Semantic Seg
%To combat this variability a lot of work has been done to perform
%automatic segmentation of liver and its tumors.
%The first methods developed to perform this task were based on
%anatomical prior-knowledge or manual interactions, and often failed due to  the limited amount of training samples, or when
%facing pathological cases. Automatic deep learning segmentation can address most of these issues by looking for relevant features directly from the images only. We review in chapter 3 both DL-based and non DL-based liver tumor segmentation methods.\\
%In chapter 4, we present our multiphase cascaded architecture to automatically segment both the liver and tumors and to partition the tumor into necrotic and active tissues.\\
%In the last chapter we present our preliminary work to automatically predict the histological grade of HCCs.
%Finally, conclusion and perspectives are given.
%We review the different \ac{dl} related liver tumors semantic segmentation studies and extract the common key settings to build our own segmentation pipeline.
%From these key settings, we focus on the cascaded architecture allowing,
%in the case of tumors segmentation, the implementation of two specialized
%networks (one to segment the liver on an entire \ac{ct} slice and the second to delineate the
%tumors, but only on the liver region).\\
%To prove the ability of \ac{dl} networks to segment liver tissues, and to
%prove that a cascaded architecture combined with multiphase CT images
%allows the best accuracy, we perform semantic segmentation of liver
%tissues on an internal dataset.\\

%We review liver radiomics studies and focus on the deep learning radiomics branch for its ability to extract features directly from the input without  necessary requiring any prior knowledge. We present our multiphase  cascaded architecture to segment the liver and
%its tumor, before applying this to weakly annotated datasets. We finally
%use extracted features from our multiphase architecture to predict the
%histological grade. \\



%To complete missing annotations in a weakly annotated or non-annotated
%dataset, we use the same architecture.




\section*{Contributions}
\label{sec:IntroContributions}
\addcontentsline{toc}{section}{\nameref{sec:IntroContributions}}

In this work we present strategies to automatically provide characterization of the liver tumors.
We present a review of current liver-related hand-crafted radiomics studies while assessing their experimental design through the Radiomics Quality Score, and propose different ways to improve their reproducibility.
Often performed manually, the delineations of the tumors can suffer from various bias, such as the inter and intra-observer variability, or its undefined borders leading to errors when diagnosing, classifying or staging the liver cancer. Therefore, we present a robust automatic pipeline to perform semantic segmentation of liver tumors that take advantage of deep cascaded networks architectures and the incorporation of dynamic contrast-enhanced information. A main prerequisite for the liver tumors characterization is the study of dynamic multiphase images, however, the major challenge that we faced in our work is the difficulty to obtain (either internally or externally) databases containing liver tumors multiphase images. To characterize the tumors, we combined several datasets to tackle the automatic histological grade prediction and we propose a novel deep radiomics pipeline which uses relevant semantic imaging features to assess the grade in a slice-wise fashion.\\


%In the first chapter we present the medical context and motivate our choices to focus on HCC with a characterization based on CT images. The second chapter present state-of-art methods to automatically segment liver and liver tumors whereas the third chapter introduce the concept of radiomics with a summary of both HCR and DLR-based existing studies applied to the liver.
%The fourth chapter present our first contribution, with a detailed description of our cascaded semantic segmentation architecture implemented to segment the liver, its tumors and to separate active and necrotic parts of tumors. In the fifth chapter, we update this architecture to propose an automatic prediction of the histological grade. Our preliminary results were compared with those obtained using a HCR-based pipeline. We concluded this manuscript by presenting several ways to improve our work.
